---
title: "4G Performance Variance Overview"
author: Networks of Guatemala
output:
  html_document:
    df_print: paged
---

<style>
body {
text-align: justify}
</style>


### What is this data?
This comes from a dataset of cell tower daily performance metrics.

```{r}
  # Set up the doc environment 
  library(knitr)
  library(ggplot2)
  library(dplyr)
  library(kableExtra)

  source("../../utils.R")
  base  <- baseDir()

  qosDataPath  <- paste(base, "data/tigo/FullPerfData4GDaily.csv", sep= '')
  cellDataPath <- paste(base, "data/tigo/cells.csv", sep= '')    

  color1 <- "#941010"
```

## Variation: 4G Performance data
Now we'll look at how quality of service (QOS) ratios vary.
QOS is calculated as `timeGoodPerf / time`

```{r}
  # import quality of service data and cell location data
  qos4g <- rio::import(qosDataPath)
  cells <- rio::import(cellDataPath)
  
  #prep dataframes for merge
  qos4g$CELL <- as.character(qos4g$CELL)
  
  # add cell location 
  qos4g <- left_join(qos4g, cells, by = c("CELL" = "CELL_ID") )
  
  # filter towers with no traffic
  qos4g <- qos4g[qos4g$TIME4G > 0, ] 
  
  # calculate performance ratios and cell-level differences from the mean
  qos4g$perfRatio4g <- qos4g$TIMENAVREQ804G / qos4g$TIME4G
  qos4g$cellPerfDiff4g <- qos4g$perfRatio4g - qos4g$CELL_AVG_PERF
  
  # calculate the global performance ratio mean
  globalMean4g <- mean(qos4g$perfRatio4g, na.rm=T)
  
  # calculate cell-level difference from global performance mean
  qos4g$globalPerfDiff4g <- qos4g$perfRatio4g - globalMean4g 
  
  # calculate variance per cell
  cellVars <- group_by(qos4g, CELL) %>% 
    summarise(cellVar = var(perfRatio4g))
  qos4g <- left_join(qos4g, cellVars)
  
``` 
```{r message= F, echo= F}
  ggplot(qos4g, aes(x=perfRatio4g)) + 
    geom_histogram(color= color1, fill= color1, binwidth=.005) + 
    labs(title="Distribution of daily cell performance", 
      x= "cell performance ratio")

  ggplot(qos4g, aes(x=cellPerfDiff4g)) + 
    geom_histogram(fill= color1, binwidth= .005) + 
    labs(title="Distribution of daily differences from cell performance mean", 
      x= "deviation from cell mean performance (daily cell perf - average cell perf)")
  
  ggplot(qos4g, aes(x= cellVar)) + 
    geom_histogram(fill= color1, binwidth= .001) + 
    labs(title="Distribution of cell-level performance ratio variance", 
      x= "cell performance ratio variance)")
```

##### Cell performance mean
```{r}
  summary(qos4g$CELL_AVG_PERF)
```
##### Mean performance vs cell variance
```{r message=F}
  #plot performance mean and average cell variance
  ggplot(qos4g, aes(x = cellVar, y = CELL_AVG_PERF)) +
    geom_point(color= color1, size=0.01) + 
    labs(x = "cell performance variance", y = "mean cell performance")

```

##### Drilling down on variance, splitting between high and low var `threshold = .05`
```{r}
  qos4g.highvar <- qos4g[qos4g$cellVar > .03,]
  qos4g.lowvar <- qos4g[qos4g$cellVar < .03,]
```
`r length(unique(qos4g.highvar$CELL)) ` cells have a variance greater than .03  

`r length(unique((qos4g.lowvar)$CELL)) ` cells have a variance less than .03

```{r}
  ggplot(qos4g.highvar, aes(x = cellVar, y = CELL_AVG_PERF)) +
    geom_point(color= color1) + 
    labs(x = "cell performance variance", y = "mean cell performance", title="High Variance")
  
  ggplot(qos4g.lowvar, aes(x = cellVar, y = CELL_AVG_PERF)) +
    geom_point(color= color1) + 
    labs(x = "cell performance variance", y = "mean cell performance", title="Low Variance")
  

```





