---
  title: "Variance in 4g cell performance and prensa access from cell"
  author: Networks of Guatemala
---

## Appendix:
```{r}
  addCellAverages <- function(prensa) {                                      
    cellAvg <- group_by(prensa, bts_sh_nm) %>%                               
      summarise(                                                             
             cellAvgPerfRatio3g = mean(perfRatio3g, na.rm = T),              
             cellAvgPerfRatio4g = mean(perfRatio4g, na.rm = T),              
             cellAvgPrensaSubs = mean(subscribers, na.rm = T), 
             cellAvgSubs = mean(subs, na.rm = T) 
      )                                                                      
                                                                             
      left_join(prensa, cellAvg) %>%                                         
        mutate(cellPerfDiff4g       = perfRatio4g - cellAvgPerfRatio4g,          
                cellPerfDiff3g      = perfRatio3g - cellAvgPerfRatio3g,          
                cellPrensaSubsDiff  = subscribers - cellAvgPrensaSubs,
                cellPrensaSubsPercentDiff   = cellPrensaSubsDiff / cellAvgPrensaSubs,
                cellSubsDiff                = subs - cellAvgSubs, 
                cellSubsPercentDiff         = cellSubsDiff / cellAvgSubs,
        )           
  }                                                                          
       

  mergeLocationData <- function(prensa, cells) {                                    
    cells$cell <- cells$CELL_ID                                              
    left_join(prensa, cells)                                                 
  }                                                                          
     

  aggregate.cellAndMonth <- function(traffic, perf) {                        
    traffic$dateAndCell <- paste(traffic$monthyear, traffic$bts_sh_nm)       
    perf$dateAndCell <- paste(perf$monthyear, perf$lvl_val)                  
                                                                             
    merge(perf, traffic, all = F)                                            
  }  

```

```{r}
  library(dplyr)
  library(knitr)
  library(ggplot2)

  qosDataPath <- "~/Guatemala/NetworksOfGuatemala/data/tigo/perfMaySept.RData"
  navDataPath <- "~/Guatemala/NetworksOfGuatemala/data/tigo/PrensaNavData.RData"
  cellDataPath <- "~/Guatemala/NetworksOfGuatemala/data/tigo/cells.csv"

  color1 <- "#941010"                                                      
```


### Data cleaning
We begin by loading our three datasets: qosData, navData, and cellData.
```{r}
  traffic <- readRDS(navDataPath)               
  perf <- readRDS(qosDataPath)               
  cells <- read.csv(cellDataPath)                                          
```


This analysis uses monthly cell-level performance and navigation data 
from May to Sept of 2018. It is meant to help answer the larger quesiton of
whether cell performance impact political engagement. 

In this analysis, we investigate whether there is a correlation between
monthly changes in a cell's performance and changes for the same month
in the number of subscribers that used this cell to access PrensaLibre.


We then add a performance ratio, which will be used later, to our performance dataset
```{r}
  perf$perfRatio3g <- perf$timenavreq3g/perf$time3g                        
  perf$perfRatio4g <- perf$timenavreq4g/perf$time4g                        
```
The `timenavreq` column gives the amount of time that this cell provided
a connection that was good for the user's activity. Therefore, this 
performance ratio represents the % of time that users which connnected 
using this cell experienced good Quality of Service.



Then we merge our network navigation data, our network cell performance data, 
and our cell information by month and cell
```{r}
  # the aggregation functions can be found in the appendix
  cellData <- aggregate.cellAndMonth(traffic, perf) %>%                              
    mergeLocationData(cells)                                                 
```

Our full cell data now looks like this
```{r}
  kable(head(cellData))
```

Now that we have our fully merged cell data, we need to perform some mutations
to create new columns that we will use for our analysis. 

We want to determine whether, in a month that a cell's performance declined,
did the number of subscribers that accessed PrensaLibre also decline. 

For this, we need the average performance and average subscriber numbers
for each cell. Then we need the amount that a cell varied from its 5-month 
mean in a given month.

We use a function `addCellAverages` to compute this and add these columns 
to our data
```{r}
    # this function can be found in the appendix
    cellData <- addCellAverages(cellData)                                                       
```

Now our data is ready for analysis. First, let's explore the data visually.


Here we'll plot the monthly variance in performance and subscriber numbers
for each (cell, month) pairing. Remember, these differences are based on
how much a particular cell in a given month varied from that same cell's 
5-month average performance ratio.
```{r echo= F, out.width="1800px"}

  ggplot(cellData, aes(x = cellPerfDiff4g, y = cellPrensaSubsDiff)) +                       
    geom_point(aes(colour = STATE), size = .1) +                                            
    labs(x = "Cell performance ratio variance", 
         y = "Cell Prensa Subscriber variance") +
    theme(
          legend.position="bottom", 
          legend.text = element_text(size=5),
          legend.direction = "horizontal",
          legend.box = "vertical"
    ) +
    guides(
           colour = guide_legend(override.aes = list(size=2)),
           fill=guide_legend(title="Department")
    )

```

One of the interesting things about this chart is the high varition if prensa
subscribers in the green departments vs the rest of the country. Let's take
a close look by building the same graph, but only looking at 
Escuintla, Guatemala, Huehuetenango, and Izabal.

```{r echo= F, out.width="1800px"}

  greenDepts <- c("ESCUINTLA", "GUATEMALA", "HUEHUETENANGO", "IZABAL")
  greenCellData <- cellData[cellData$STATE %in% greenDepts,]

  ggplot(greenCellData, aes(x = cellPerfDiff4g, y = cellPrensaSubsDiff)) +                       
    geom_point(aes(colour = STATE), size = .1) +                                            
    labs(x = "Cell performance ratio variance", 
         y = "Cell Prensa Subscriber variance") +
    theme(
          legend.position="bottom", 
          legend.text = element_text(size=5),
          legend.direction = "horizontal",
          legend.box = "vertical"
    ) +
    guides(
           colour = guide_legend(override.aes = list(size=2)),
           fill=guide_legend(title="Department")
    )

```
This view indicates that the variance is coming from Guatemala. This makes
sense since PrensaLibre is primarily read here in the capital. 


Let's check this assumption by looking at the variance by municipality in the
department of Guatemala.
```{r echo= F, out.width="1800px"}

  guateCellData <- cellData[cellData$STATE == "GUATEMALA",]

  ggplot(guateCellData, aes(x = cellPerfDiff4g, y = cellPrensaSubsDiff)) +                       
    geom_point(aes(colour = CITY), size = .1) +                                            
    labs(x = "Cell performance ratio variance", 
         y = "Cell Prensa Subscriber variance") +
    theme(
          legend.position="bottom", 
          legend.text = element_text(size=5),
          legend.direction = "horizontal",
          legend.box = "vertical"
    ) +
    guides(
           colour = guide_legend(override.aes = list(size=2)),
           fill=guide_legend(title="Department")
    )

```
As expected, the majority of variation in prensa readers is coming from
the capital.

Now we'll look at the variance in the capital grouped by month
```{r echo= F, out.width="1800px"}
  
  capitalCellData  <- guateCellData[guateCellData$CITY == "GUATEMALA",]

  ggplot(capitalCellData, aes(x = cellPerfDiff4g, y = cellPrensaSubsDiff)) +                       
    geom_point(aes(colour = monthyear), size = .1) +                                            
    labs(x = "Cell performance ratio variance", 
         y = "Cell Prensa Subscriber variance") +
    theme(
          legend.position="bottom", 
          legend.text = element_text(size=5),
          legend.direction = "horizontal",
          legend.box = "vertical"
    ) +
    guides(
           colour = guide_legend(override.aes = list(size=2)),
           fill=guide_legend(title="Department")
    )

```



Now we'll build a linear regression model to examine the correlation
```{r}
    summary(lm(cellData$cellPrensaSubsDiff ~ cellData$cellPerfDiff4g))    
```
This model indicates a positive correlation between
monthly variation in the number of subscribers accessing prensa and
the monthly variation in cell performance. However, the $R^2$ value is 
extremely small, indicating that a small amount of the variation is 
explained by our dependent variable. 

As we showed in our graphs above, 
most of the variation in prensa subscribers comes from the capital. 
Therefore, let's run the same analysis focused on Guatemala City.

```{r}
    capitalCellData <- guateCellData[guateCellData$CITY == "GUATEMALA",]
    ols.model = lm(capitalCellData$cellPrensaSubsDiff ~ capitalCellData$cellPerfDiff4g)    
    summary(ols.model)
```
As expected, this result increases the correlation and $R^2$ value - though
it it remains extremely small. The coefficient displayed here indicates 
that a 1% deviation from the average performance for a given cell is 
correlated with a shift of 2.46 subscribers around the average number of 
prensa subscribers from a given cell.


For a better understanding of this coefficient, we'll produce a confidence interval.
```{r}
  confint(ols.model)
```
Building off of our initial logic, this indicates that a 1% shift in a cells
performance ratio from it's mean performance is correlated with a shift 
in the number of subscribers that access Prensa Libre of between 
2.07 and 2.85.



Let's put this into context. How many subscribers are usually accessing 
prensa from the cells in the capital?
```{r}
  summary(capitalCellData$subscribers)
```
With a mean of 25.23 subscribers and a median of 7.00 subscribers, these
results indicate that a 1% change in cell performance is correlated with
an 8.2% shift in prensa subscribers in the worst case and a 40.7% in the best
case vs. the global average.

###Overall subscriber shifts

Now that we've seen the correlation between Prensa Subscribers and 
cell performance, what about total subscribers? We'll look at this here.

Variance across the country
```{r echo= F, out.width="1800px"}

  ggplot(cellData, aes(x = cellPerfDiff4g, y = cellSubsDiff)) +                       
    geom_point(aes(colour = STATE), size = .1) +                                            
    labs(x = "Cell performance ratio variance", 
         y = "Cell Prensa Subscriber variance") +
    theme(
          legend.position="bottom", 
          legend.text = element_text(size=5),
          legend.direction = "horizontal",
          legend.box = "vertical"
    ) +
    guides(
           colour = guide_legend(override.aes = list(size=2)),
           fill=guide_legend(title="Department")
    )

```

Variance in the same departments we filtered for before 
```{r echo= F, out.width="1800px"}

  ggplot(greenCellData, aes(x = cellPerfDiff4g, y = cellSubsDiff)) +                       
    geom_point(aes(colour = STATE), size = .1) +                                            
    labs(x = "Cell performance ratio variance", 
         y = "Cell Prensa Subscriber variance") +
    theme(
          legend.position="bottom", 
          legend.text = element_text(size=5),
          legend.direction = "horizontal",
          legend.box = "vertical"
    ) +
    guides(
           colour = guide_legend(override.aes = list(size=2)),
           fill=guide_legend(title="Department")
    )

```

Variance throughout Guatemala
```{r echo= F, out.width="1800px"}
  ggplot(guateCellData, aes(x = cellPerfDiff4g, y = cellSubsDiff)) +                       
    geom_point(aes(colour = CITY), size = .1) +                                            
    labs(x = "Cell performance ratio variance", 
         y = "Cell Subscriber variance") +
    theme(
          legend.position="bottom", 
          legend.text = element_text(size=5),
          legend.direction = "horizontal",
          legend.box = "vertical"
    ) +
    guides(
           colour = guide_legend(override.aes = list(size=2)),
           fill=guide_legend(title="Department")
    )

```

Variance throughout Guatemala
```{r echo= F, out.width="1800px"}
  ggplot(capitalCellData, aes(x = cellPerfDiff4g, y = cellSubsDiff)) +                       
    geom_point(aes(colour = monthyear), size = .1) +                                            
    labs(x = "Cell performance ratio variance", 
         y = "Cell Subscriber variance") +
    theme(
          legend.position="bottom", 
          legend.text = element_text(size=5),
          legend.direction = "horizontal",
          legend.box = "vertical"
    ) +
    guides(
           colour = guide_legend(override.aes = list(size=2)),
           fill=guide_legend(title="Department")
    )

```


Now we'll build a model to examine the correlation
```{r}
    summary(lm(cellData$cellSubsDiff ~ cellData$cellPerfDiff4g))    
```
This model indicates a positive correlation between
monthly variation in the number of subscribers accessing prensa and
the monthly variation in cell performance. However, the $R^2$ value is 
extremely small, indicating that a small amount of the variation is 
explained by our dependent variable. 

As we showed in our graphs above, 
most of the variation in prensa subscribers comes from the capital. 
Therefore, let's run the same analysis focused on Guatemala City.

```{r}
    capitalCellData <- guateCellData[guateCellData$CITY == "GUATEMALA",]
    ols.model = lm(capitalCellData$cellSubsDiff ~ capitalCellData$cellPerfDiff4g)    
```






###Correlation or Causation

*TODO: Need to build an alaysis to answer the questions below*
Now we ask whether this correlation is causal. The biggest area for error
arises from how we measure performance. To measure performance we use the 
number of minutes in which a call provided good performance given the activity
of the user. It is possible that increased traffic to Prensa shifted the 
distribution of content accessed towards text and away from heavier activity
like video streaming. In this case, the increased Prensa traffic would 
cause the performance deviation.



















