---
  title: "Variance in 4g cell performance and prensa access from cells in the capital"
  author: Networks of Guatemala
---

```{r include= F}
  ## Appendix:
  addCellAverages <- function(prensa) {                                      
    cellAvg <- group_by(prensa, bts_sh_nm) %>%                               
      summarise(                                                             
             cellAvgPerfRatio3g = mean(perfRatio3g, na.rm = T),              
             cellAvgPerfRatio4g = mean(perfRatio4g, na.rm = T),              
             cellAvgPrensaSubs = mean(subscribers, na.rm = T), 
             cellAvgSubs = mean(subs, na.rm = T) 
      )                                                                      
                                                                             
      left_join(prensa, cellAvg) %>%                                         
        mutate(cellPerfDiff4g       = perfRatio4g - cellAvgPerfRatio4g,          
                cellPerfDiff3g      = perfRatio3g - cellAvgPerfRatio3g,          
                cellPrensaSubsDiff  = subscribers - cellAvgPrensaSubs,
                cellPrensaSubsPercentDiff   = cellPrensaSubsDiff / cellAvgPrensaSubs,
                cellSubsDiff                = subs - cellAvgSubs, 
                cellSubsPercentDiff         = cellSubsDiff / cellAvgSubs,
        )           
  }                                                                          
       

  mergeLocationData <- function(prensa, cells) {                                    
    cells$cell <- cells$CELL_ID                                              
    left_join(prensa, cells)                                                 
  }                                                                          
     

  aggregate.cellAndMonth <- function(traffic, perf) {                        
    traffic$dateAndCell <- paste(traffic$monthyear, traffic$bts_sh_nm)       
    perf$dateAndCell <- paste(perf$monthyear, perf$lvl_val)                  
                                                                             
    merge(perf, traffic, all = F)                                            
  }  

```

```{r include= F}
  library(dplyr)
  library(knitr)
  library(ggplot2)

  qosDataPath <- "~/Guatemala/NetworksOfGuatemala/data/tigo/perfMaySept.RData"
  navDataPath <- "~/Guatemala/NetworksOfGuatemala/data/tigo/PrensaNavData.RData"
  cellDataPath <- "~/Guatemala/NetworksOfGuatemala/data/tigo/cells.csv"

  color1 <- "#941010"                                                      
```

###Overview
This analysis uses monthly cell-level performance and navigation data 
from May to Sept of 2018. It is meant to help answer the larger quesiton of
whether cell performance impacts political engagement. 

In this analysis, we investigate whether there is a correlation between
monthly changes in a cell's performance and changes for the same month
in the % of the number of subscribers that used this cell to access 
PrensaLibre.


###Data cleaning
We begin by loading our three datasets: qosData, navData, and cellData.
```{r}
  traffic <- readRDS(navDataPath)               
  perf <- readRDS(qosDataPath)               
  cells <- read.csv(cellDataPath)                                          
```


We then add a performance ratio, which will be used later, to our
performance dataset
```{r}
  perf$perfRatio3g <- perf$timenavreq3g/perf$time3g                        
  perf$perfRatio4g <- perf$timenavreq4g/perf$time4g                        
```
The `timenavreq` column gives the amount of time that this cell provided
a connection that was good for the user's activity. Therefore, this 
performance ratio represents the % of time that users which connnected 
using this cell experienced good Quality of Service.



Then we merge our network navigation data, our network cell performance data, 
and our cell information by month and cell
```{r}
  cellData <- aggregate.cellAndMonth(traffic, perf) %>%                              
    mergeLocationData(cells)                                                 
```


Now that we have our fully merged cell data, we need to perform some mutations
to create new columns that we will use for our analysis. 

We want to determine whether, in a month that a cell's performance declined,
did the number of subscribers that accessed PrensaLibre also decline. 

For this, we need the average performance and average subscriber numbers
for each cell. Then we need the amount that a cell varied from its 5-month 
mean in a given month and the % change of prensa subscribers vs the 5 month cell mean.

We use a function `addCellAverages` to compute these and add the columns 
to our data. In addition, we filter our data to focus on the capital.
```{r message= F}
    cellData <- addCellAverages(cellData)                                                       
    guateCellData <- cellData[cellData$STATE == "GUATEMALA",]
    capitalCellData <- guateCellData[guateCellData$CITY == "GUATEMALA",]
```

Now our data is ready for analysis. First, let's explore the data visually.

******************************************************
###Analysis
Here we'll plot the monthly variance in performance and subscriber numbers
for each (cell, month) pairing. Remember, these differences are based on
how much a particular cell in a given month varied from that same cell's 
5-month average performance ratio.
```{r echo= F, out.width="1800px"}
  ggplot(capitalCellData, aes(x = cellPerfDiff4g, y = cellPrensaSubsPercentDiff)) +                       
    geom_point(color= color1, size = .1) +                                            
    labs(x = "Cell performance ratio variation", 
         y = "Cell Prensa Subscriber % variation")

```

Now we'll look at the variance in the capital grouped by month
```{r echo= F, out.width="1800px"}
  
  capitalCellData  <- guateCellData[guateCellData$CITY == "GUATEMALA",]

  ggplot(capitalCellData, aes(x = cellPerfDiff4g, y = cellPrensaSubsPercentDiff)) +                       
    geom_point(aes(colour = monthyear), size = .1) +                                            
    labs(x = "Cell performance ratio variance", 
         y = "Cell Prensa Subscriber % variation") +
    theme(
          legend.position="bottom", 
          legend.text = element_text(size=10),
          legend.direction = "horizontal",
          legend.box = "vertical"
    ) +
    guides(
           colour = guide_legend(override.aes = list(size=2)),
           fill=guide_legend(title="Department")
    )

```
Performance data seems to be grouped by month. 
We'll take a closer look at this later.


Now we'll build a linear regression model to examine the correlation
```{r}
    ols.model = lm(capitalCellData$cellPrensaSubsPercentDiff ~ capitalCellData$cellPerfDiff4g)    
    summary(ols.model)
```


For a better understanding of this coefficient, we'll produce a confidence interval.
```{r}
  confint(ols.model)
```
This indicates that a 1% shift in a cells
performance ratio from it's mean performance is correlated with a shift 
in the number of subscribers that access Prensa Libre of between 
6.25% and 7.3%.

************************************************************

###Performance by month
We notices a relationship between performance and month before.
Here we'll take a closer look at this correlation with a linear regression
of performance differences by month.
```{r}
    ols.model = lm(capitalCellData$cellPerfDiff4g ~ capitalCellData$monthyear)    
    summary(ols.model)
    confint(ols.model)
```
This model confirms our hunch. We see a 10% average performance difference
from May to June. This is a *large* monthly difference for a city of 
3 million people.



Now let's look at this same model but for cells across the country.
```{r}
    ols.model = lm(cellData$cellPerfDiff4g ~ cellData$monthyear)    
    summary(ols.model)
    confint(ols.model)
```
Though not quite as pronounced, we still see drastic shifts in performance 
by month. Again, changes of this magnitude are surprising for a 10 million 
user cell network.

************************************************************

###Total subscriber shifts
Now that we've seen the correlation between Prensa Subscribers and 
cell performance, what about total subscribers? We'll look at this here.

Variance across the country
```{r echo= F, out.width="1800px"}

  ggplot(cellData, aes(x = cellPerfDiff4g, y = cellSubsPercentDiff)) +                       
    geom_point(aes(colour = STATE), size = .1) +                                            
    labs(x = "Cell performance ratio variance", 
         y = "Cell Subscriber % shift") +
    theme(
          legend.position="bottom", 
          legend.text = element_text(size=5),
          legend.direction = "horizontal",
          legend.box = "vertical"
    ) +
    guides(
           colour = guide_legend(override.aes = list(size=2)),
           fill=guide_legend(title="Department")
    )

```


Variance throughout the capital
```{r echo= F, out.width="1800px"}
  ggplot(capitalCellData, aes(x = cellPerfDiff4g, y = cellSubsPercentDiff)) +                       
    geom_point(aes(colour = monthyear), size = .1) +                                            
    labs(x = "Cell performance ratio variance", 
         y = "Cell Subscriber % shift") +
    theme(
          legend.position="bottom", 
          legend.text = element_text(size=10),
          legend.direction = "horizontal",
          legend.box = "vertical"
    ) +
    guides(
           colour = guide_legend(override.aes = list(size=2)),
           fill=guide_legend(title="Department")
    )

```

Now we'll build a model to examine the correlation between overall 
subscriber % shifts and performance changes in the capital
```{r}
    summary(lm(capitalCellData$cellSubsPercentDiff ~ capitalCellData$cellPerfDiff4g))    
```
Here we see a correlation that indicates a 0.3% shift in subscriber 
numbers for a 1% change in local cell performance. This 0.3% for all subscribers
is compared to a 6.8% change in subscribers to Prensa Libre specifically.

This indicates that navigation to Prensa Libre is more sensative to local
performance shifts than overall traffic and supports the thinking that
increased network performance disproportionately increased consumption of
political information.


******************************************
###Other question
What if our inference here is inverted? If people read prensa instead of 
watching netflix, the threshold for what would be considered good 
performance would shift because lower performance is required to read text.

In this scenario and increase in Prensa traffic could an increase in
our performance measures.

We're confident that this is not the case because traffic to PrensaLibre.com
is not a large enough portion of total traffic to change overall 
performance numbers.


###Next steps
This is promising as a preliminary result; however, it is far from 
conclusive.

To expand on this analysis our next steps are to look at all departments
and analyze news consumption using a set of websites, rather than just one.

Furthermore, by using more granular levels of data aggregation 
(daily or weekly), we will be able to dramatically increase the power of 
the analysis.



