---
title: "3G Performance Variance Overview"
author: Networks of Guatemala
output:
  html_document:
    df_print: paged
---

```{r Setup environment and load data, cache=TRUE}
  # Set up the doc environment 
  library(ggplot2)
  library(dplyr)

  source("../../utils.R")
  base  <- baseDir()
  color1 <- "#941010"
  
  qosData3gPath  <- paste(base, "data/tigo/QosDailyNational3g.csv", sep= '')
  qos3g <- rio::import(qosData3gPath)
```
## Variation: 3G Performance data
For each cell, the telco records the amount of time that users experienced a good connection for the activity given
the activity they are performing. Note that this 'good performance' threshold is different for reading articles
and streaming video. 

The performance of a given cell is calculated as `good performance time / total connection time` which 
produces a performance ratio. This is the ratio we use to measure performance.

```{r Add performance metrics}
  # add performance ratio numbers
  qos3g$perfRatio3g <- qos3g$TIMENAVREQ803G / qos3g$TIME3G 

  globalMean3g <- mean(qos3g$perfRatio3g, na.rm=T)
  print(globalMean3g)

  # calculate avg performance ratios over time per cell (shortId).
  # If a cell is replaced because it breaks it could have a different 
  # longId but may take on the shortId of the cell it replaces.
  library(dplyr)
  cellAvg <- group_by(qos3g, CELL) %>%
  summarise(
      avgPerf3g = mean(perfRatio3g),
      varPerf3g = var(perfRatio3g))

  # merge cell performance averages with the rest of the dataset and
  # calculate deviation from cell performance mean
  qos3g <- left_join(qos3g, cellAvg)
``` 
```{r Plot performance ratios}
  ggplot(qos3g, aes(x=avgPerf3g)) + 
    geom_histogram(color= color1, fill= color1, binwidth=.005) + 
    labs(title="3g average cell performance", 
      x= "mean cell performance ratio")
  
  ggplot(qos3g, aes(x=varPerf3g)) + 
    geom_histogram(color= color1, fill= color1, binwidth= .005) + 
    labs(title="Distribution of deviations from cell performance mean", 
      x= "deviation from cell mean performance")
```

##### Cell performance variance and averages
```{r Summary and variance outputs}
  summary(qos3g$avgPerf3g)
  var(qos3g$avgPerf3g)
  var(qos3g$perfRatio3g)
```
##### Mean performance vs cell variance
```{r Plot mean performance vs. variance}
  ggplot(qos3g, aes(x = varPerf3g, y = avgPerf3g)) +
    geom_point(color= color1) + 
    labs(x = "cell performance variance", y = "mean cell performance")

```

##### Drilling down on variance, splitting between high and low var `threshold = .05`
```{r Plot high and low variance cells}
  qos3g.highvar <- qos3g[qos3g$varPerf3g > .05,]
  qos3g.lowvar <- qos3g[qos3g$varPerf3g < .05,]
  
  ggplot(qos3g.highvar, aes(x = varPerf3g, y = avgPerf3g)) +
    geom_point(color= color1) + 
    labs(x = "cell performance variance", y = "mean cell performance", title="High Variance")
  
  ggplot(qos3g.lowvar, aes(x = varPerf3g, y = avgPerf3g)) +
    geom_point(color= color1) + 
    labs(x = "cell performance variance", y = "mean cell performance", title="Low Variance")
```





